---
title: "Assignment 1"
subtitle: "STATS 220 Semester One 2022"
author: "Nev Yildirim"
date: "21/03/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(magick)
```

## GitHub repository information
* Please visit [my STATS220 repo](https://github.com/neverkam/stats220) to find the incredible work we do in STATS220 - Data Technologies paper. 

## Animated GIF creation
```{r}
gif_image <- image_read("https://st2.depositphotos.com/1173077/8707/v/950/depositphotos_87073942-stock-illustration-boo-ghost-cartoon.jpg") %>% image_scale(500)

frame_1 <- gif_image %>%
  image_scale(200) %>%
  image_extent("500x500")

frame_2 <- gif_image %>%
  image_scale(300) %>%
  image_extent("500x500")

frame_3 <- gif_image %>%
  image_scale(400) %>%
  image_extent("500x500")

frame_4 <- gif_image %>%
  image_scale(500) 

# putting the frames in a vector and creating an animation
gif <- c(frame_1, frame_2, frame_3, frame_4) %>% image_animate(fps = 1)
gif

```
```{css}
@import url('https://fonts.googleapis.com/css2?family=Montserrat:wght@100&family=Palette+Mosaic&display=swap');

h2 {font-family: 'Red Hat Mono', monospace;}
body {
  font-family: 'Montserrat', sans-serif;
} 
```

## Learning reflection
What a start to the year! I love the fact that I could not only create memes but also moving memes (in other words gifs) and I love how simple it actually was when taught properly. I was always uncertain about how to use vectos in R. In these past two weeks, that uncertainty was taken care of and I am much more comfortable with vector operations now. 

As I was confused about vectors, I was also quite confused about data frames. So I am definitely looking forward to working with data frames.


