---
title: "Assignment 5"
author: "Nev Yildirim"
date: "01/06/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, error = FALSE)
library(tidyverse)
library(ggplot2)
library(gganimate)
library(jsonlite)
```

## Introduction
I have decided to create a data frame that I combined using the _Apple Music Top 100_ playlists from four different countries. These are **New Zealand**, **UK**, **USA**, **South Africa**. My purpose was to analyse the music trends in the English speaking world(common-wealth countries). Although I couldn't compare and contrast there countries' listening trends, I have found their most favourite artists.
## Visualisation

```{r}
library(tidyverse)
library(ggplot2)
library(gganimate)

song_data <- readRDS("song_data.rds")

song_data_separated <- song_data %>%
  separate_rows(artistName, sep = ", | & " ) %>%
  group_by(artistName) %>%
  summarise(num_songs = n()) %>%
  arrange(desc(num_songs)) %>%
  filter(num_songs >= 10)

song_data_separated %>% ggplot() +
  geom_col(aes(y=artistName, x=num_songs), fill = "#6a0dad") +
  labs(x = "Number of songs", y = "Artist name",
       title ="Top artists in all 4 playlists",
       subtitle = "Who are the most favourite artists in all these 4 countries?",
       caption = "Source: Apple Music") +
  theme(axis.text.y = element_text(angle = 45)) +
  theme(panel.background = element_rect(fill =
                                          "white"))

```


In this data analysis, I am trying to figure out the artist names who are being listened to to a larger extent than most other artists in each of the playlists of each country. In addition to that, I have filtered the data in a way that we can only look at the artists who have more than 10 songs in these playlists. As a result, we can clearly observe that **Harry Styles** seems to be the most favourite artist at this stage, with a total of over approximately 60 songs in all the four playlists.  

## Data creation

## Final reflection
I had been looking for Module 5 since the day STATS220 started, as I was always curious about web scraping. Web scraping was such a topic that both intrigued me and made me nervous. Inspecting web pages and _borrowing_ information from them sounded absolutely incredible and promising, but when I saw the confusing and lengthy HTML code, I felt uncomfortable when I first came across **web scraping**. However, after Module 5, I realised there is a patter to things and when I figure out that pattern, specifically to be able to know which part of the HTML code you're wanting to scrape, things get much easier. 

Overall, I am looking forward to working with SQL more, maybe take my knowledge and look into the design aspects of SQL databases and join data using my R knowledge.

## Code appendix

```{r file='data_web.R', eval=FALSE}
library(tidyverse)
library(rvest)
library(jsonlite)
library(gganimate)

urls <- c("https://music.apple.com/us/playlist/top-100-new-%20zealand/pl.d8742df90f43402ba5e708eefd6d949a",
          "https://music.apple.com/us/playlist/top-100-uk/pl.c2273b7e89b44121b3093f67228918e7",
          "https://music.apple.com/us/playlist/top-100-usa/pl.606afcbb70264d2eb2b51d8dbcfa6a12",
          "https://music.apple.com/za/playlist/top-100-south-africa/pl.447bd05172824b89bd745628f7f54c18")

apple_data <- map_df(1 : length(urls), function(i){
  
  # taking two seconds break
  Sys.sleep(2)
  page <- read_html(urls[i])
  
  trackId <- page %>%
    html_elements(".songs-list") %>%
    html_elements("a") %>%
    html_attr("href") %>%
    .[str_detect(.,"song")] %>%
    str_remove_all("https://music.apple.com/us/song/") %>%
    parse_number()
  
  track_name <- page %>%
    html_elements(".songs-list-row__song-name") %>%
    html_text2()
  
  track_length <- page %>%
    html_elements(".songs-list-row__length") %>%
    html_text2()
  
  track_length_min <- 
    60 * str_sub(track_length, 1, 1) %>% parse_number()
  
  track_length_secs <- 
    str_sub(track_length, 3, 4) %>% parse_number()
  
  track_length_total <- track_length_min + track_length_secs
  
  
  return(tibble(trackId, track_name, track_length_total))
})

apple_data

saveRDS(apple_data, "apple_data.rds")

```

```{r file='data_api.R', eval=FALSE}
library(tidyverse)
library(jsonlite)
readRDS("apple_data.rds")

trackIds <- apple_data$trackId %>% unique() %>% na.omit() %>% parse_number()
trackIds

itunes_data <- map_df(1 : length(trackIds), function(i){
  Sys.sleep(2)
  song_id <- trackIds[i]
  url <- paste0("https://itunes.apple.com/lookup?id=", song_id)
  response <- fromJSON(url)
  itunes_data <- response$results
})

saveRDS(itunes_data, "itunes_data.rds")


```

```{r file='data_join.R', eval=FALSE}
library(tidyverse)
apple_data <- readRDS("apple_data.rds")
itunes_data <- readRDS("itunes_data.rds")
song_data <- inner_join(apple_data, itunes_data, by = 'trackId')
saveRDS(song_data, "song_data.rds")

```

```{r file='data_viz.R', eval=FALSE}
library(tidyverse)
library(ggplot2)
song_data <- readRDS("song_data.rds")

song_data_separated <- song_data %>%
  separate_rows(artistName, sep = ", | & " ) %>%
  group_by(artistName) %>%
  summarise(num_songs = n()) %>%
  arrange(desc(num_songs)) %>%
  filter(num_songs >= 10)

song_data_separated %>% ggplot() +
  geom_col(aes(y=artistName, x=num_songs), fill = "#6a0dad") +
  labs(x = "Number of songs", y = "Artist name",
       title ="Top artists in all 4 playlists",
       subtitle = "Who are the most favourite artists in all these 4 countries?",
       caption = "Source: Apple Music") +
  theme(axis.text.y = element_text(angle = 45)) +
  theme(panel.background = element_rect(fill =
                                          "white"))
```

```{r file='data_creativity.R', eval=FALSE}
library(tidyverse)
library(ggplot2)
library(gganimate)

song_data <- readRDS("song_data.rds")

song_data_separated <- song_data %>%
  separate_rows(artistName, sep = ", | & " ) %>%
  group_by(artistName) %>%
  summarise(num_songs = n()) %>%
  arrange(desc(num_songs)) %>%
  filter(num_songs >= 10)

song_data_separated %>% ggplot() +
  geom_col(aes(y=artistName, x=num_songs), fill = "#6a0dad") +
  labs(x = "Number of songs", y = "Artist name",
       title ="Top artists in all 4 playlists",
       subtitle = "Who are the most favourite artists in all these 4 countries?",
       caption = "Source: Apple Music") +
  theme(axis.text.y = element_text(angle = 45)) +
  theme(panel.background = element_rect(fill =
                                          "white"))
```

```{css}
@import url('https://fonts.googleapis.com/css2?family=Palette+Mosaic&family=Red+Hat+Mono:wght@300&family=Roboto:wght@300&display=swap');

h2 {font-family: 'Red Hat Mono', monospace;}
body {
  font-family: 'Roboto', sans-serif;
} 
```

